---
title: "plant traits database"
output: html_document
--- 

**CREATING THE FINAL SPECIES LIST**
This .Rmd file contains code that combines the list of species from three different sites at Danum: the 50ha site (unlogged), the CP site (logged) and the INDFORSUS site (both logged and unlogged). 

```{r}
library(here)#for telling R where my working directory is
library(dplyr)#for manipulating dataframes
library(stringr)#for working with strings
```

note: I couldn't install BIEN because my R would not install the package RPostgreSQL. This code seemed to fix the problem: install.packages("RPostgreSQL", type = "binary")

**CREATING THE DANUM SPECIES LIST**

THE 50ha SITE (UNLOGGED)
```{r}
#read in the list of unlogged tree species from the 50ha site
unlogged_species <- read.csv(here("species list data", "Danum_spp_list_unlogged.csv"))

#create a new column with the full binomial latin name of each species
unlogged_species$binomial_name <- paste(unlogged_species$Genus, unlogged_species$Species)

#create a dataframe of just the species names of unlogged trees.
unlogged_species_list <- unlogged_species %>% 
  select(binomial_name, Family, Genus, Species) %>% #selects just the columns we want
  mutate(logging_status = "unlogged") %>% #adds a new column called logging_status, and fills it with 'unlogged'
  mutate(species_origin = "50ha", logging_status = "unlogged") %>% #adds a column called species_origin, and fills it with '50ha'
  distinct(binomial_name, .keep_all=TRUE) #removes any duplicates from the list, based only on binomial_name

```


THE LOGGED SITE (CP PLOTS)
```{r}
#read in the dataset of logged tree species from the CP site
logged_species <- read.csv(here("species list data", "DanumRevised_35Plots_FullPlotData_Trees_2020_edit_4.11.csv"))

#now I need to read in the data on the restoration status of the CP sites
logged_restoration_status <- read.csv(here("species list data", "LoggedForest_PlotList_UpdatedMar2024.csv"))


#create a column with the full binomial latin name of each species
logged_species$binomial_name <- paste(logged_species$Genus, logged_species$Species)

#join logged_species with the restoration status of each species, and the select the columns I am interested in 
logged_species <- left_join(logged_species, logged_restoration_status, by = "PlotNo") %>%
  select(PlotNo, restoration, binomial_name, Family, Genus, Species, DBH.mm, Height_m) %>%
  filter(binomial_name != "NA NA") %>%
  mutate(species_origin = "CP")

#before going any further, save this file as a .csv, so that it can be used to calculate relative abundance in a separate .Rmd
write.csv(logged_species, file = here("species list data", "CP plots logged species.csv"), row.names = FALSE)


```

```{r}
#make a list of distinct species names from the CP plot.
#this dataset also has info on DBH and height, do I maybe want to keep this info?
logged_species_list <- logged_species %>%
  select(binomial_name, Family, Genus, Species, restoration) %>% 
  mutate(logging_status = "logged") %>%
  mutate(species_origin = "CP") %>%
  group_by(binomial_name, Family, Genus, Species, species_origin, logging_status) %>%
  summarise(
    restoration = str_c(unique(restoration), collapse = ", "),
    .groups = "drop")

#make the Family listed in title case, rather than in all capital letters.  
logged_species_list$Family <- str_to_title(str_to_lower(logged_species_list$Family))

```


```{r}
#now combine the logged_species_list and the unlogged_species_list together
#bind_rows() combines the two dataframes together
#group_by() groups according to binomial name
#summarize() - if a species appears more than once in the list, it is assigned "both" as its logging status, otherwise, it is assigned the logging_status that it already had when the dataframes were separate
#ungroup() ungroups the data

# Bind the two dataframes
combined_species <- bind_rows(logged_species_list, unlogged_species_list) %>%
  group_by(binomial_name, Genus, Species) %>%
  summarise(
    Family = first(Family),
    logging_status = ifelse(n_distinct(logging_status, na.rm = TRUE) > 1, "both", first(logging_status)),
    species_origin = str_c(unique(species_origin[!is.na(species_origin)]), collapse = ", "),
    .groups = "drop") 

combined_species_list <- combined_species %>%
  left_join(
    logged_species_list %>%
      select(binomial_name, restoration) %>%
      distinct(),
    by = "binomial_name") %>%
  rename(history_at_CP_site = restoration)

```


THE INDFORSUS SITE
```{r}
#adding in lists of species from actively restored or naturally regenerated areas
#read in the data of the species found in the INFORSUS (logged) site
INDFORSUS_data <- read.csv(here("species list data", "mature_data_RH(in).csv"))

#read in the data on their restoration status
INDFORSUS_restoration_status_data <- read.csv(here("species list data", "2018_2019Seedling_RH(Plot_Info).csv"))

#now join the two datasets together while keeping information on dbh, height, and wood density
#INDFORSUS_species_list <- INDFORSUS_data %>%
  #left_join(INDFORSUS_restoration_status_data, by = "station", relationship = "many-to-many") %>% # join the two datasets together according to their station
  #select(history, diameter_mm, height_m, wooddensity, family, genus, species) %>% # select only the columns that we are interested in
  #distinct() %>% # remove rows that have the same information in every single column
  #mutate(binomial_name = paste(genus, species, sep = " ")) %>% # add in a column with the binomial_name
  #mutate(species_origin = case_when(
    #grepl("ActiveRestoration|NaturalRegeneration", history) ~ "IF",
    #grepl("Unlogged", history) ~ "50ha",
    #TRUE ~ NA_character_ # default if none of the conditions match
#)) %>%
  #mutate(logging_status = case_when(
    #history %in% c("ActiveRestoration", "NaturalRegeneration") ~ "logged",
    #history == "Unlogged" ~ "unlogged",
    #TRUE ~ NA_character_)) %>% #handels other cases; not really necessary here
  #rename( #change the names of the columns so that they are written with capital letters
    #Family = family,
    #Genus = genus,
    #Species = species)


#now join the two datasets together keeping only the list of species.
INDFORSUS_species_list <- INDFORSUS_data %>%
  left_join(INDFORSUS_restoration_status_data, by = "station", relationship = "many-to-many") %>% # join the two datasets together according to their station
  select(history, family, genus, species) %>% # select only the columns that we are interested in
  distinct() %>% # remove rows that have the same information in every single column
  mutate(binomial_name = paste(genus, species, sep = " ")) %>% # add in a column with the binomial_name
  mutate(species_origin = case_when(
    grepl("ActiveRestoration|NaturalRegeneration", history) ~ "IF",
    grepl("Unlogged", history) ~ "50ha",
    TRUE ~ NA_character_ # default if none of the conditions match
  )) %>%
  mutate(logging_status = case_when(
    history %in% c("ActiveRestoration", "NaturalRegeneration") ~ "logged",
    history == "Unlogged" ~ "unlogged",
    TRUE ~ NA_character_)) %>% #handels other cases; not really necessary here
  rename( #change the names of the columns so that they are written with capital letters
    Family = family,
    Genus = genus,
    Species = species)


INDFORSUS_species_list <- INDFORSUS_species_list %>%
  filter(species_origin != "50ha") %>% #remove all rows where species_origin = 50ha, so that I am not duplicating information
  rename(history_at_IF_site = history) %>%
  group_by(binomial_name, Family, Genus, Species) %>%
  summarise(
    species_origin = str_c(sort(unique(species_origin[!is.na(species_origin)])), collapse = ", "),
    history_at_IF_site = str_c(sort(unique(history_at_IF_site[!is.na(history_at_IF_site)])), collapse = ", "),
    .groups = "drop"
  )


```


```{r}
#now I want to join the  final_species list and the INDFORSUS species lists together, according to the binomial_name column.
#
#final_species_list <- bind_rows(combined_species_list, INDFORSUS_species_list) %>%
  #group_by(binomial_name, Genus, Species) %>%
  #summarize(
    #Family = first(Family),
   # species_origin = toString(sort(unique(species_origin))),#makes the species origin column
    #logging_status = if (n_distinct(logging_status) > 1) "both" else first(logging_status),
    #.groups = "drop") %>%
#  filter(!is.na(Family)) %>% #removes the last row which is full of NAs. specifically, where the family is NA.
#clean up the history column
  ##mutate(history = str_replace_all(history, "\\bNA\\b", "")) %>%  # Replace 'NA' with empty string
  #mutate(history = str_replace_all(history, "^,|,$", "")) %>%     # Remove leading/trailing commas
  #mutate(history = ifelse(history == "", NA, str_trim(history))) %>%
  #rename(history_at_INDFORSUS_site = history)# Trim spaces and set empty to NA


#create the final species list
final_species_list <- bind_rows(combined_species_list, INDFORSUS_species_list) %>%
  group_by(binomial_name, Genus, Species) %>%
  summarise(
    Family = first(Family),
    history_at_CP_site = first(history_at_CP_site),
    species_origin = str_c(sort(unique(species_origin[!is.na(species_origin)])), collapse = ", "),
    logging_status = if (n_distinct(logging_status[!is.na(logging_status)]) == 1) {
      first(logging_status[!is.na(logging_status)])
    } else {
      "both"
    },
    .groups = "drop"
  )

#add on the IF history, because for some reason i couldn't do this within the pipe above
IF_history <- INDFORSUS_species_list %>%
  select(binomial_name, history_at_IF_site)

final_species_list <- left_join(final_species_list, IF_history, by = "binomial_name")

#save the final_species_list as a .csv file
write.csv(final_species_list, file = here("species list data", "final species list.csv"), row.names = FALSE)
 
```



