---
title: "sta"
author: "Ugo Soggiu"
date: '2024-03-15'
---
```{r message=FALSE, warning=FALSE}
install.packages("BIEN")
library(BIEN)
library(readr)
library(tidyverse)
```

# test bain package 

```{r}
Osuri_regen_data <- read_delim("Osuri_regen_data.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
Osuri_tree_data <- read_delim("Osuri_tree_data.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
tree_old <- read_delim("tree_olddata.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
Bahamas_species_list <- BIEN_list_country(country = "Bahamas")
India_species_list <- BIEN_list_country(country = "India")
trait_list = BIEN_trait_list()

Osuri_species_adult_2 = tree_old[!duplicated(tree_old$species), ]

Osuri_species_seedl = Osuri_regen_data[!duplicated(Osuri_regen_data$Species_name), ]
Osuri_species_adult = Osuri_tree_data[!duplicated(Osuri_tree_data$Species_name), ]
Osuri_species_list = as.character(Osuri_species_adult$Species_name)

# Find species in species_table1 that are not in species_table2
species_not_in_table2 <- setdiff(Osuri_species_adult$Species_name, Osuri_species_adult_2$species)

# Find species in species_table2 that are not in species_table1
species_not_in_table1 <- setdiff(Osuri_species_adult_2$species,Osuri_species_adult$Species_name)

# Find species common to both tables
common_species <- intersect(Osuri_species_adult_2$species,Osuri_species_adult$Species_name)

all_species <- union(union(species_not_in_table2, species_not_in_table1), common_species)

#write.csv(all_species, file = "species_list.csv", row.names = FALSE)

```

# Species list with seedling 

```{r}
Osuri_seedl_data <- read_delim("Osuri_anamalais_data/Osuri_regen_data.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
Osuri_species_seedl = Osuri_regen_data[!duplicated(Osuri_regen_data$Species_name), ]

Regen_old <- read_delim("Osuri_anamalais_data/regen_olddata.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
Old_species_seedl = Regen_old[!duplicated(Regen_old$Species.name), ]

# Find species in species_table1 that are not in species_table2
species_1 <- setdiff(Old_species_seedl$Species.name, Osuri_species_seedl$Species_name)

# Find species in species_table2 that are not in species_table1
species_2 <- setdiff(Osuri_species_seedl$Species_name,Old_species_seedl$Species.name)

# Find species common to both tables
common_species_2 <- intersect(Old_species_seedl$Species.name,Osuri_species_seedl$Species_name)

all_species_2 <- union(union(species_2, species_1), common_species_2)

diff_adult_seed_1 <- setdiff(all_species,all_species_2)
diff_adult_seed_2 <- setdiff(all_species_2,all_species)
common_species_to <- intersect(all_species_2,all_species)
common_species_old <- intersect(Old_species_seedl$Species.name,Osuri_species_adult_2$species)
common_species_new <- intersect(Osuri_species_seedl$Species_name,Osuri_species_adult$Species_name)

all_species_tot <- union(all_species, all_species_2)

write.csv(all_species_tot, file = "species_list_tot.csv", row.names = FALSE)
```

## synonym species updated
The species list was cross checked with World Plant Online : https://powo.science.kew.org/
Synonyms and error was corrected and noted via excel. 
Here we apply this correction to the presence-absence table 

```{r}
Osuri_regen_data <- read_delim("Osuri_regen_data.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
list_species = read_delim("species_list_tot.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
for(i in 1 : 1632){
  for(j in 1 : 207){
    
  if (!is.na(list_species[j, 3])){
    if (Osuri_regen_data[i,4] == list_species[j,3]){
    Osuri_regen_data[i,4] <- list_species[j,1]
  } 
  }
  if (!is.na(list_species[j, 4])){
    if(Osuri_regen_data[i,4] == list_species[j,4]){
    Osuri_regen_data[i,4] <- list_species[j,1]
  }
  }
  }
}
Osuri_tree_data <- read_delim("Osuri_tree_data.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
for(i in 1 : 2195){
  for(j in 1 : 207){
    
  if (!is.na(list_species[j, 3])){
    if (Osuri_tree_data[i,4] == list_species[j,3]){
    Osuri_tree_data[i,4] <- list_species[j,1]
  } 
  }
  if (!is.na(list_species[j, 4])){
    if(Osuri_tree_data[i,4] == list_species[j,4]){
    Osuri_tree_data[i,4] <- list_species[j,1]
  }
  }
  }
}
tree_old <- read_delim("tree_olddata.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
for(i in 1 : 3430){
  for(j in 1 : 207){
    
  if (!is.na(list_species[j, 3])){
    if (tree_old[i,10] == list_species[j,3]){
    tree_old[i,10] <- list_species[j,1]
  } 
  }
  if (!is.na(list_species[j, 4])){
    if(tree_old[i,10] == list_species[j,4]){
    tree_old[i,10] <- list_species[j,1]
  }
  }
  }
}
old_regen <- read_delim("old_regen.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
for(i in 1 : 600){
  for(j in 1 : 207){
    
  if (!is.na(list_species[j, 3])){
    if (old_regen[i,8] == list_species[j,3]){
    old_regen[i,8] <- list_species[j,1]
  } 
  }
  if (!is.na(list_species[j, 4])){
    if(old_regen[i,8] == list_species[j,4]){
    old_regen[i,8] <- list_species[j,1]
  }
  }
  }
}
#write.csv(tree_old, file = "old_tree_cor.csv", row.names = FALSE)
#write.csv(Osuri_tree_data, file = "OS_tree_cor.csv", row.names = FALSE)
#write.csv(Osuri_regen_data, file ="OS_regen_cor.csv", row.names = FALSE)
#write.csv(old_regen, file ="old_regen_cor.csv", row.names = FALSE)
```

#Trait available_Osuri 
One leaf trait dataset from Osuri study of 2024 was shared with us. 
This dataset was crossed with the species lists of study site to create a table of trait. 
The first operation was to check for which species trait data are available ->

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Tree_trait <- read_delim("IN_TRAITS.MASTER.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
Species_ID <- read_delim("species.master.tplfixed.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
list_species = read_delim("species_list_tot.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)

#remove all the duplica
Species_trait = Tree_trait[!duplicated(Tree_trait$species), ]

# Utilisez match() pour obtenir les indices des noms complets des espèces correspondant aux codes d'espèces dans Tree_trait$species
matching_indices <- match(Species_trait$species, Species_ID$spcode)

# Utilisez les indices pour obtenir les noms complets des espèces correspondant aux codes d'espèces dans Tree_trait$species
full_species_names <- Species_ID$species[matching_indices]

# Ajoutez la nouvelle colonne contenant les noms complets des espèces à Tree_trait
Species_trait$full_species_name <- full_species_names

## we do the same with our own species list
# Utilisez match() pour obtenir les indices des noms complets des espèces correspondant aux codes d'espèces dans Tree_trait$species
matching_indices <- match(Species_trait$species, list_species$sp_code)

# Utilisez les indices pour obtenir les noms complets des espèces correspondant aux codes d'espèces dans Tree_trait$species
full_species_names <- list_species$species[matching_indices]

# Ajoutez la nouvelle colonne contenant les noms complets des espèces à Tree_trait
Species_trait$plot_species_name <- full_species_names

# Reorganise col. 
colonne_a_deplacer = Species_trait[,c(32,33)]
Species_trait <- Species_trait[, -c(32,33)] #remove extra col
Species_trait <- cbind(Species_trait[, 1:8], colonne_a_deplacer, Species_trait[, 9:ncol(Species_trait)]) #Insert col in the right place

Trait_available = na.omit(Species_trait$plot_species_name)
Missing_species = list_species[!(list_species$species %in% Trait_available), ]
#write.csv(Missing_species, file = "Missing_species.csv", row.names = FALSE)

```

#Fruit table (finish all trait good ! )
Another datasets with the fruit and dispersal trait. 
Same operation. 

```{r}
Fruit_data <- read_delim("Anamalais_trait_data_2024.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)

Fruit_table = as.data.frame(list_species[,c(1,2)])

matching_indices <- match(list_species$species,Fruit_data$Accept_name_WFO)
full_species_names <- Fruit_data[matching_indices, 5:16]
Fruit_table[,3:14] <- full_species_names[,1:12]
Fruit_table[c(41,44,90,206),3:14] <- Fruit_data[c(49,53,118,261),5:16]
#write.csv(Fruit_table, file = "Fruit_table.csv", row.names = FALSE)
```

## Seed Bien 
```{r}
Dispers_data <- read_delim("Fruit_table.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
List_sp = Dispers_data$species
#list of availible trait in BIEN 
trait_list = BIEN_trait_list()

Seed_mass_BIEN = BIEN_trait_mean(species = List_sp, trait = "seed mass")
Seed_length_BIEN = BIEN_trait_mean(species = List_sp, trait = "seed length")
#write.csv(Seed_mass_BIEN, file = "Seed_mass_Bien.csv", row.names = FALSE)
```

#Creation of the final trait table
First, calculate the median trait value per species. 
Then, cross with the species list from our plot. 

```{r}

Tree_trait = Tree_trait[,-c(1:8,28:31)]
Species_Osu = Tree_trait[!duplicated(Tree_trait$species), ]

#calculation of the mean trait value (Osuri data)

Mean_trait <- as.data.frame(matrix(NA, nrow = 245, ncol = 19)) #creation of empty matrix to welcome the mean trait value (per species)
colnames(Mean_trait) = colnames(Tree_trait)
Mean_trait[,1] = Species_Osu$species
sample = list()
i = 1
j =2
for(i in 1:245){                    #loop which calculate the mean trait value for each trait per species
  y = as.character(Mean_trait[i,1])
  Species = as.data.frame(subset(Tree_trait, species == y))
  sample[i] = nrow(Species)              # number of samples per species 
  for(j in 2:19){
   x = mean(Species[,j], na.rm = TRUE) #why mean and not median ? 
   Mean_trait[i,j] <- x                
  }
}
Mean_trait$sample = sample

#alternative without loop 

# Calculate the mean of each col per species
#Mean_Species <- Tree_trait_trans %>%
 # group_by(plot_species_name) %>%
  #summarise_all(~ mean(., na.rm = TRUE)) %>%
  #rename_with(~ paste0("mean_", .), -plot_species_name)

# Link "Mean_Species" with "Mean_trait"
#Mean_trait <- left_join(Mean_trait, Mean_Species, by = c("species" = "plot_species_name"))


# Final table creation
Leaf_table = as.data.frame(list_species[,c(1,2)])

matching_indices <- match(list_species$sp_code,Mean_trait$species) 
full_species_names <- Mean_trait[matching_indices, 2:20]
Leaf_table[,3:21] <- full_species_names[,1:19]
Leaf_table_Na <- na.omit(Leaf_table[,c(1:3)])   #Just a test to see the number of species with available trait data 
# Sélectionner les colonnes avec au moins une valeur NA
Leaf_missng <- anti_join(Leaf_table, Leaf_table_Na)
Leaf_table$sample <- sapply(Leaf_table$sample, function(x) ifelse(length(x) == 1, x, NA))

#write.csv(Leaf_table, file = "Leaf_table_OS.csv", row.names = FALSE)
```

# Bien trait filling the gaps 
To fill the data gaps, we have used the Database BIEN from the package BIEN. 
Here, the different datasets are selected from the package. 

```{r}
Leaf_Genus <- read_delim("Leaf_table_genus.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
Leaf_Genus_Na <- na.omit(Leaf_Genus[,c(1:3)])   #Just a test to see the number of species with available trait data 
# Sélectionner les colonnes avec au moins une valeur NA
Genus_missng <- anti_join(Leaf_Genus, Leaf_Genus_Na)

List_missing = as.character(as.list(Leaf_missng[,1]))

#list of availible trait in BIEN 
trait_list = BIEN_trait_list()

#LDMC datasets
LDMC_BIEN = BIEN_trait_mean(species = List_missing, trait = "leaf dry mass per leaf fresh mass")
#SLA datasets 
SLA_BIEN = BIEN_trait_mean(species = List_missing, trait = "leaf area per leaf dry mass")
# LeafN (nitrogen)
LeafN_BIEN = BIEN_trait_mean(species = List_missing, trait = "leaf nitrogen content per leaf dry mass")
# LeafP (phosphorus)
LeafP_BIEN = BIEN_trait_mean(species = List_missing, trait = "leaf phosphorus content per leaf dry mass")
# Seed Size
Seed_size_BIEN = BIEN_trait_mean(species = List_missing, trait = "seed length")
# Height
#Height_BIEN = BIEN_trait_mean(species = List_missing, trait = "whole plant height")

occurrences <- BIEN_occurrence_species(species = List_missing)

#write.csv(LDMC_BIEN, file = "LDMC_BIEN.csv", row.names = FALSE)
#write.csv(SLA_BIEN, file = "SLA_BIEN.csv", row.names = FALSE)
#write.csv(LeafN_BIEN, file = "LeafN_BIEN.csv", row.names = FALSE)
#write.csv(Seed_size_BIEN, file = "Seed_size_BIEN.csv", row.names = FALSE)
#write.csv(LeafP_BIEN, file = "LeafP_BIEN.csv", row.names = FALSE)
```

# Final crossing value
Crossing between BIEN and Osuri Data
```{r}
Leaf_Genus <- read_delim("Leaf_table_genus.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
Leaf_table <- read_delim("Leaf_table_OS.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
#remove unnecessary col
Trait_genus = Leaf_Genus[,-c(6:11,18:21)]
Trait_table = Leaf_table[,-c(5:10,17:20)]
Trait_table$Level = Leaf_Genus$Level
#add sources 
Trait_table$Source <- ifelse(is.na(Trait_table$Level), "BIEN", "OS")
#add genus level 
Trait_table[,c(3:11)] = Trait_genus[,c(4:12)]

## BIEN DATA 
SLA_BIEN <- read_delim("SLA_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
LDMC_BIEN <- read_delim("LDMC_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
SLA_genus = SLA_BIEN[which(SLA_BIEN$level_used != "Family"),]

#put SLA BIEN value 
Trait <- Trait_table %>%
  left_join(SLA_genus[,c(1,5,6)], by = "species") %>%
  mutate(sla_cm2.g = ifelse(is.na(sla_cm2.g), `cm²/g`, sla_cm2.g)) %>%
  select(-c(`cm²/g`))
#LDMC
## unit not correct for the OSURI DATA, transform in mg/g
LDMC_genus = LDMC_BIEN[which(LDMC_BIEN$level_used != "Family"),]
Trait <- Trait %>% 
  mutate(LDMC= LDMC * 1000) %>% #mg/g
  left_join(LDMC_genus[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(LDMC = ifelse(is.na(LDMC), mean_value, LDMC)) %>%
  select(-c(mean_value))
#Nitrogen 
N_BIEN <- read_delim("LeafN_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
N_genus = N_BIEN[which(N_BIEN$level_used != "Family"),]
Trait <- Trait %>% 
  mutate(leafN = leafN * 10) %>% #mg/g
  left_join(N_genus[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(leafN = ifelse(is.na(leafN), mean_value, leafN)) %>%
  select(-c(mean_value))
#Phosphorus 
P_BIEN <- read_delim("LeafP_BIEN.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
P_genus = P_BIEN[which(P_BIEN$level_used != "Family"),]
Trait <- Trait %>% 
  mutate(leafP = leafP * 10) %>% #mg/g
  left_join(P_genus[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(leafP = ifelse(is.na(leafP), mean_value, leafP)) %>%
  select(-c(mean_value))
#write.csv(Trait, file = "Trait.csv", row.names = FALSE)
```

## add Family 
The data was add in three step : the genus level from Osuri - genus from BEIN - Family from BIEN 
The Family from Osuri was to short. Here, it's the final step. 
```{r}
## BIEN DATA 
SLA_BIEN <- read_delim("SLA_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
LDMC_BIEN <- read_delim("LDMC_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
N_BIEN <- read_delim("LeafN_BIEN.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
P_BIEN <- read_delim("LeafP_BIEN.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
Trait <- read_delim("Trait.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
#SLA
count_SLA <- sum(is.na(Trait$sla_cm2.g))
##SLA BIEN value 
Trait <- Trait %>%
  left_join(SLA_BIEN[,c(1,5,6)], by = "species") %>%
  mutate(sla_cm2.g = ifelse(is.na(sla_cm2.g), `cm²/g`, sla_cm2.g)) %>%
  select(-c(`cm²/g`))
#LDMC 
count_LDMC <- sum(is.na(Trait[,7]))
#add
Trait <- Trait %>% 
  left_join(LDMC_BIEN[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(LDMC = ifelse(is.na(LDMC), mean_value, LDMC)) %>%
  select(-c(mean_value))
#N
count_N <- sum(is.na(Trait[,9]))
#add
Trait <- Trait %>% 
  left_join(N_BIEN[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(leafN = ifelse(is.na(leafN), mean_value, leafN)) %>%
  select(-c(mean_value))
#P
count_P <- sum(is.na(Trait[,10]))
#add
Trait <- Trait %>% 
  left_join(P_BIEN[,c(1,5,2)], by = "species") %>% #same thing that for SLA 
  mutate(leafP = ifelse(is.na(leafP), mean_value, leafP)) %>%
  select(-c(mean_value))
#write.csv(Trait, file = "Trait_final.csv", row.names = FALSE)
```


##WD and height
For Wood density (WD) and height, the presence dataset of tree and seedling contain WD and height value. 
So theses values are collected in the final table. 
```{r}
Osuri_regen_data <- read_delim("OS_regen_cor.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
Osuri_tree_data <- read_delim("OS_tree_cor.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
tree_old <- read_delim("old_tree_cor.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)
Trait <- read_delim("Trait_final.csv", delim = ",", escape_double = FALSE, trim_ws = TRUE)

WD_Species = Osuri_tree_data[,-c(1:3,6:18,29:31)]

#Mean of each species 
sp = Osuri_tree_data[!duplicated(Osuri_tree_data$Species_name),-c(1:3,6:18,29:31) ]
Mean_trait <- as.data.frame(matrix(NA, nrow = 130, ncol = 12)) #creation of empty matrix to welcome the mean trait value (per species)
colnames(Mean_trait) = colnames(WD_Species)
Mean_trait[,c(1,3:6,9:11)] = sp[,c(1,3:6,9:11)]
sample = list()
ine = c(2,7,8,12)
for(i in 1:133){                    #loop which calculate the mean trait value for each trait per species
  y = as.character(Mean_trait[i,1])
  Species = as.data.frame(subset(WD_Species, Species_name == y))
  sample[i] = nrow(Species)              # number of samples per species 
  for(j in ine){
   x = mean(Species[,j], na.rm = TRUE) #why mean and not median ? 
   Mean_trait[i,j] <- x                
  }
}
Mean_trait$sample = sample
WD_Species = Mean_trait
colnames(WD_Species)[colnames(WD_Species) == "Species_name"] <- "species"
WD_Species = as.data.frame(WD_Species)

# Left join on 'species'
Trait_t <- Trait[,-c(14:23)] %>%
  left_join(WD_Species[, c("species", "Wden_final")], by = "species")

# Update 'wd_g.cm3' with 'Wden_final' for common species
Trait_t <- Trait_t %>%
  mutate(wd_g.cm3 = ifelse(!is.na(Wden_final), Wden_final, wd_g.cm3)) %>%
  select(-Wden_final)  # Remove 'Wden_final' after updating

#Same for ad_ht
Trait_t <- Trait_t %>%
  left_join(WD_Species[,c(1,2)], by = "species") %>%
  mutate(ht_m = ifelse(is.na(ht_m), Height, ht_m)) %>%
  select(-c(Height))

            ## Same operation with seedling ##
WD_Species = Osuri_regen_data[,c(4,11,15)]

#Mean of each species 
sp = Osuri_regen_data[!duplicated(Osuri_regen_data$Species_name),c(4,11,15) ]
Mean_trait <- as.data.frame(matrix(NA, nrow = 150, ncol = 3)) #creation of empty matrix to welcome the mean trait value (per species)
Mean_trait[,1] = sp[,1]
colnames(Mean_trait) = colnames(WD_Species)
sample = list()
i =1
for(i in 1:150){                    #loop which calculate the mean trait value for each trait per species
  y = as.character(sp[i,1])
  Species = as.data.frame(subset(WD_Species, Species_name == y))
  sample[i] = nrow(Species)              # number of samples per species 
  for(j in 2:3){
   x = mean(Species[,j], na.rm = TRUE) #why mean and not median ? 
   Mean_trait[i,j] <- x                
  }
}
Mean_trait$sample = sample
WD_Species = Mean_trait
colnames(WD_Species)[colnames(WD_Species) == "Species_name"] <- "species"
WD_Species = as.data.frame(WD_Species)

# Left join on 'species'
Trait_t <- Trait_t %>%
  left_join(WD_Species[,c(1,2)], by = "species") %>%
  mutate(wd_g.cm3 = ifelse(is.na(wd_g.cm3), Wden_final, wd_g.cm3)) %>%
  select(-c(Wden_final))

#Same for ad_ht
Trait_t <- Trait_t %>%
  left_join(WD_Species[,c(1,3)], by = "species") %>%
  mutate(ht_m = ifelse(is.na(ht_m), ad_ht, ht_m)) %>%
  select(-c(ad_ht))

Trait_t[Trait_t == "NaN"] <- NA

#write.csv(Trait_t, file = "Trait_final.csv", row.names = FALSE)
```




